A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-07-08 bioc-workflow.Rmd
Added TL reference

Using the functions `setAnnotationParams` and `makeGoSet` we can contruct an auxiliary `MSnSet` of GO terms, from the primary data's features i.e. the protein accession numbers. All the GO terms associated to each accession number are retrieved and used to create a binary matrix where a 1 (0) at position (i; j) indicates that term j has (not) been used to annotate protein i. The GO terms are retrieved from an appropriate repository using the `r Biocpkg("biomaRt")` package. The specific `r Biocpkg("Biomart")` repository and query will depend on the species under study and the type of identifiers. The first step is to condtruct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`. Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study (below, left), then what type of identifier is used to annotate the proteins in your `MSnSet` (below, right).

![]()

It is also possible to pass patterns to match the species e.g. "Mus musculus", and the identifier used e.g. "Uniprot/Swissprot Accession", for the `Biocpkg("Biomart")` query.

```{r setap}
par <- setAnnotationParams(inputs = c("Mus musculus", 
                                      "Uniprot/Swissprot Accession"))

@@

```{r}
gocc <- makeGoSet(hl, params = par, 
                  namespace = "cellular_component", 
                  evidence = NULL)
gocc
```

Now we have generated our auxiliary data we can use the $k$-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` for the predictions.

The $k$-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.

One useful source of information that may help us solve our task is Gene Ontology information, such as the cellular compartment namespace. We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with a higher generalisation accuracy than using standard supervised machine learning with a single source of information [@Breckels:2016].

In the code chunk below, we extract information from the Gene Ontology to use as an auxiliary data source to be intergrated with our high quality MS-based data.

>>>>

Using the functions `setAnnotationParams` and `makeGoSet` we can contruct an auxiliary `MSnSet` of GO terms, from the primary data's features i.e. the protein accession numbers. All the GO terms associated to each accession number are retrieved and used to create a binary matrix where a 1 (0) at position (i; j) indicates that term j has (not) been used to annotate protein i. The GO terms are retrieved from an appropriate repository using the `r Biocpkg("biomaRt")` package. The specific `r Biocpkg("Biomart")` repository and query will depend on the species under study and the type of identifiers. The first step is to condtruct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`. Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study, and then what type of identifier is used to annotate the proteins in your `MSnSet`. It is also possible to pass patterns to match the species e.g. in the code chunk below we pass "Mus musculus", and the identifier used e.g. "Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.

```{r setap}
par <- setAnnotationParams(inputs = c("Mus musculus", 
                                      "UniProt/Swissprot"))

@@

```{r makego}
gocc <- makeGoSet(hl, params = par, 
                  namespace = "cellular_component", 
                  evidence = NULL)
gocc
```

Now we have generated our auxiliary data we can use the *k*-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` to do the predictions. We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with a higher generalisation accuracy than using standard supervised machine learning with a single source of information [@Breckels:2016].

The *k*-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.
